We presented a novel approach to specify distributed dataflows.
It is based on the contracts, preconditions and postconditions of the graph operations.
We considered a small machine leaning streaming pipeline example, discussed ways of its optimization and saw that our approach defines the most optimal execution graph too.
% TODO We demonstrated the prototype?
Also we outlined optimization challenges that we faced.
Finally, we showed ideas for the future work enhancements.

We proposed to specify dataflow as CGraph (contracted graph).
CGraph is a pair of environment and semantics, where environment maps nodes to their contracts, and semantics is a set of nodes that produce dataflow target side-effects.
We provide an algorithm that generates all concrete execution graphs that satisfy contracts and semantics.

We discussed the example that SQL implementation seems to be unnatural.
And the only possible way to specify it is execution graph, that represents rather imperative approach then declarative, hence do not provide enough space for optimizations.
We specified the example problem by CGraph and pointed out that the most optimal graph exists in the set of concrete execution graphs, generated from CGraph.
% TODO Way of reasoning about problems by contracts in declarative way

We pointed out the optimization problems that we face.
% TODO Preliminary cost evaluation
% TODO Runtime statistics gathering
% TODO Cost evaluation by statistics analysis
% TODO Actor entity that decides, whether to restructure execution graph
% TODO How to restructure execution graph in runtime

In future TODO
% TODO General implementation
% TODO Better contracts interface (trello)
% TODO Smart advices
