% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

% listings
\usepackage{xcolor,listings}
\usepackage{textcomp}
\lstset{upquote=true}

\begin{document}
%
\title{Calco: Contract-based Approach\\ to Specify Dataflows}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Andrey Stoyan\inst{1} \and
Artem Trofimov\inst{1} \and
Nikita Sokolov\inst{1} \and
Boris Novikov\inst{2} \and
Igor E. Kuralenok\inst{1}}
%
\authorrunning{A. Stoyan et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Yandex \and
National Research University Higher School of Economics}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    Distributed dataflows can be specified in two ways: by defining an execution graph or specifying the desired result, e.g., using SQL.
    An execution graph consists of arbitrary operations, so it cannot be automatically transformed for optimization.
    SQL is popular for data analytics, and its optimization is well-researched.
    However, it is sometimes inconvenient or impossible to use SQL for general data management tasks, such as ETL, machine learning pipelines, etc.
    In this work, we introduce a novel approach to specify distributed dataflows.
    Our method is based on declarative specifications of user-defined operations called contracts.
    Such specifications allow us to build a set of equivalent graphs, which form a space for optimization.
    We implement a graphs generation prototype and outline the challenges regarding the optimization problem.

\keywords{Optimization.}
\end{abstract}
%
%
%
\section{Introduction}
\input{intro.tex}

\section{Running example}
\input{example.tex}

\section{Calco overview}
\input{overview.tex}

\section{Implementation and challenges}
\input{eval.tex}

\section{Conclusion}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{bibliography/flame-stream}

\end{document}
