\documentclass[sigconf]{acmart}

\usepackage{graphicx}
\usepackage{algorithm} % for algorithms
\usepackage{algpseudocode}
\usepackage{booktabs} % For formal tables
\usepackage{amsthm} % For claims

\theoremstyle{remark}

\settopmatter{printacmref=false, printccs=true, printfolios=true}
\pagestyle{empty} % removes running headers

\newcommand{\PicScale}{0.5}
\newcommand {\FlameStream} {FlameStream}
\begin{document}

% \copyrightyear{2018} 
% \acmYear{2018} 
% \setcopyright{acmcopyright}
% \acmConference[BeyondMR'18]{Algorithms and Systems for MapReduce and Beyond }{June 15, 2018}{Houston, TX, USA}
% \acmBooktitle{BeyondMR'18: Algorithms and Systems for MapReduce and Beyond , June 15, 2018, Houston, TX, USA}
% \acmPrice{15.00}
% \acmDOI{10.1145/3206333.3209273}
% \acmISBN{978-1-4503-5703-6/18/06}

\title {Poster: decentralized validation of statistical properties in distributed stream processing}

% \author{  Igor E. Kuralenok,$^1$     Artem Trofimov,$^ {1,2}$    Artem Permiakov,$^ {2}$   Andrey Pirozhkov,$^ {2}$    Semen Kutuzov,$^ {2}$     Artur,$^ {2}$    and  Boris Novikov$^ {1,2}$ }
% \affiliation{%
% \institution{$^1$JetBrains Research}
%   \city{St. Petersburg} 
%   \country{Russia}
% }
% \affiliation{%
% \institution{$^2$Saint Petersburg State University}
%   \city{St. Petersburg} 
%   \country{Russia}
% }
% \email{\string{ikuralenok, trofimov9artem\string}@gmail.com, borisnov@acm.org}

\begin{abstract}
Streaming applications often require validation of statistical properties. The satisfaction of such properties can directly influence the correctness of the final results. For example, in A/B testing task, there is a need to ensure that users are uniformly distributed among application versions. Otherwise, experiment results may lead to inaccurate business decisions. There are a lot of statistical tools that aim to solve this problem. Most of them assume that the whole data is processed on a single computational unit. However, this approach does not fit in distributed stream processing, because merging all data to a single machine limits throughput and causes extra communication cost. In this work, we investigate the decentralized method for statistical data validation of data streams: each computational unit independently verifies data that is assigned to it. We demonstrate statistical limitations and challenges that we recognized during experiments on a uniform and discrete data. We also show that the API of state-of-the-art stream processing systems is not enough to solve this problem efficiently. Possible solutions to the mentioned issues and open problems are discussed.      
\end {abstract}

% \keywords{Data streams, exactly-once, drifting state, optimistic OOP}

\maketitle

\thispagestyle{empty}

\section {Introduction}
\input {fs-short-intro.tex}

\section {Decentralized statistical validation}
\input {fs-short-model.tex}

\section {Experiments}
\input {fs-short-experiments.tex}

\section {Conclusion}
\input {fs-short-conclusion.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../bibliography/flame-stream}

\end {document}

\endinput
