\documentclass[sigconf]{acmart}

\usepackage{graphicx}
\usepackage{algorithm} % for algorithms
\usepackage{algpseudocode}
\usepackage{booktabs} % For formal tables
\usepackage{amsthm} % For claims
\usepackage{bbm} % indicator function
\usepackage{acmcopyright}


% table
\usepackage[flushleft]{threeparttable} % http://ctan.org/pkg/threeparttable
\usepackage{booktabs,caption}

\theoremstyle{remark}

\settopmatter{printacmref=false, printccs=true, printfolios=true}
\pagestyle{empty} % removes running headers

\newcommand{\PicScale}{0.5}
\newcommand {\FlameStream} {FlameStream}
\begin{document}

% \copyrightyear{2019} 
% \acmYear{2019} 
% \setcopyright{rightsretained} 
% \acmConference[DEBS '19]{DEBS '19: The 13th ACM International Conference on Distributed and Event-based Systems}{June 24--28, 2019}{Darmstadt, Germany}
% \acmBooktitle{DEBS '19: The 13th ACM International Conference on Distributed and Event-based Systems (DEBS '19), June 24--28, 2019, Darmstadt, Germany}\acmDOI{10.1145/3328905.3332514}
% \acmISBN{978-1-4503-6794-3/19/06}

\title {Concept Drift Aware Classification of Text Streams}

\author{Artem Trofimov,$^ {1,2}$    Mikhail Shavkunov,$^3$    Sergey Reznick,$^4$     Nikita Sokolov,$^{5}$   Mikhail Yutman,$^3$ \\   Igor E. Kuralenok,$^1$    and  Boris Novikov$^ {3}$ }
\affiliation{%
\institution{$^1$JetBrains Research}
}
\affiliation{%
\institution{$^2$Saint Petersburg State University}
}
\affiliation{%
\institution{$^3$National Research University Higher School of Economics}
}
\affiliation{%
\institution{$^4$ Kofax}
}
\affiliation{%
\institution{$^5$ ITMO University}
  \city{St. Petersburg} 
  \country{Russia}
}
\email{\string{trofimov9artem, mv.shavkunov, sergey.reznick, faucct, myutman, ikuralenok\string}@gmail.com, borisnov@acm.org}

\begin{abstract}

Text stream classification is a hot research topic with many industrial applications. However, typical approaches to static texts classification are hardly suitable for streaming setup. One of the main difficulties regarding this task is a {\em concept drift}: the changes of statistical properties of an input stream over time. According to a text classification problem, such changes include a shift of term meanings, appearance, and vanishing of target variables, etc. These changes may significantly influence the performance of a machine learning model if it is not constantly adjusting to the ongoing changes. In this work, we demonstrate the effects of concept drift on a stream of real news stories. We show that a pre-trained offline machine learning model is not able to correctly classify news topics after significant world events occur. On the other hand, it is indicated that windowed {\em IDF} features together with online learning algorithms can mitigate the issue and provide classification results which are much closer to the real news topics distribution.

\end{abstract}

% \begin{CCSXML}
% \begin{CCSXML}
% \begin{CCSXML}
% <ccs2012>
% <concept>
% <concept_id>10002951.10002952.10002953.10010820.10003208</concept_id>
% <concept_desc>Information systems~Data streams</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
% <concept>
% <concept_id>10002951.10003317.10003347.10003356</concept_id>
% <concept_desc>Information systems~Clustering and classification</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
% <concept>
% <concept_id>10002951.10003227.10003351.10003446</concept_id>
% <concept_desc>Information systems~Data stream mining</concept_desc>
% <concept_significance>300</concept_significance>
% </concept>
% </ccs2012>
% \end{CCSXML}

% \ccsdesc[500]{Information systems~Data streams}
% \ccsdesc[500]{Information systems~Clustering and classification}
% \ccsdesc[300]{Information systems~Data stream mining}

% \keywords{Data streams, text classification, reproducibility, exactly once}

\maketitle

\thispagestyle{empty}

\section {Introduction}
\input {fs-classification-intro.tex}

\section {Background}
\input {fs-classification-bgr.tex}

\section {Demonstration}
\input {fs-classification-demo.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../bibliography/flame-stream}

\end {document}

\endinput
