\documentclass[sigconf]{acmart}
% packages should be added as needed 
\usepackage{graphicx}

\usepackage{algorithm} % for algorithms
\usepackage{algpseudocode}

\usepackage{booktabs} % For formal tables

\usepackage{amsthm} % For claims
\theoremstyle{remark}

\settopmatter{printacmref=false, printccs=false, printfolios=false}
\pagestyle{plain} % removes running headers

\newcommand{\PicScale}{0.5}
\newcommand {\FlameStream} {FlameStream}
\begin{document}

\title {\FlameStream: Model and Runtime for Distributed Analytical Stream Processing}
% \titlenote{Produces the permission block, and   copyright information}
%\subtitle{Extended Abstract}
% \subtitlenote{The full version of the author's guide is available as   \texttt{acmart.pdf} document}

% \author {Igor E. Kuralenok \and Nikita Marshalkin \and Artem Trofimov \and Boris Novikov}

\author{Igor E. Kuralenok}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{JetBrains Research}
%streetaddress{P.O. Box 1212}
  \city{St. Petersburg} 
%  \state{Ohio} 
%  \postcode{190000}
  \country{Russia}
}
\email{ikuralenok@gmail.com}

\author{Artem Trofimov}
\affiliation{%
  \institution{JetBrains Research}
  %streetaddress{P.O. Box 1212}
  \city{St. Petersburg} 
  \country{Russia}
}
\email{trofimov9artem@gmail.com}

\author{Nikita Marshalkin}
\affiliation{%
\institution{JetBrains Research}
  %streetaddress{P.O. Box 1212}
  \city{St. Petersburg} 
  \country{Russia}
}
\email{marnikitta@gmail.com}

\author{Boris Novikov}
\affiliation{%
\institution{JetBrains Research}
  %streetaddress{P.O. Box 1212}
  \city{St. Petersburg} 
  \country{Russia}
}
\email{borisnov@acm.org}


%\author{Artem Trofimov, Nikita Marshalkin, Boris Novikov}
%  \authornote{}
%\affiliation{%
%  \institution{St. Petersburg state University}
%  \streetaddress{Universitetskaya nab. 7/9}
%  \city{St. Petersburg} 
%  \postcode{199034}
% \country{Russia}
%}
%\email{trofimov9artem@gmail.com, marnikitta@gmail.com, borisnov@acm.org}


\begin{abstract}
In recent years, there has been a growth in research and industrial solutions in the field of distributed stream processing. However, even state-of-the-art stream processing systems are inconvenient to work with in some ways. Firstly, existing solutions suppose that the state of operations should be managed directly by a user, and that can be confusingly for inexperienced users of analytical processing systems. Secondly, these systems do not provide for deterministic processing by default. Furthermore, the most common approach for this feature is buffering. Its main problem is the extra cost for blocking before each order-sensitive operation. The goal of this paper is to propose a model, which addresses both recognized issues. On the one hand, our model stateless from the business logic perspective. Such behavior is obtained by the technique that allows a state to be a part of a stream. On the other hand, it is deterministic by design. We introduce an optimistic method that avoids buffering before each operation to achieve determinism with low overhead. The experiments show that prototype of our system requires a low extra cost for supporting proposed features and is able to outperform alternative industrial solutions in certain conditions.
\end {abstract}

\maketitle

\section {Introduction}
\input {fs-run-time-intro.tex}

\section {Computational model}
\input {fs-run-time-model.tex}

\section{Drifting state}
\input {fs-run-time-drifting.tex}

\section {Deterministic computations}
\input {fs-run-time-deterministic.tex}

\section {Implementation}
\input {fs-run-time-impl.tex}

\section {Experiments}
\input {fs-run-time-experiments.tex}

\section {Related work}
\input {fs-run-time-related.tex}

\section {Conclusion and future work}
\input {fs-run-time-conclu.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../bibliography/flame-stream}
\end {document}


\endinput
you can put whatever here
