Most of the material in this dissertation is derived from work that has already been published and peer-reviewed. Particularly, Chapter~\ref{thesis-chapter-delivery-guarantees} and Chapter~\ref{thesis-chapther-optimistic} include content from papers on optimistic technique for achieving determinism and the exactly-once approach on top of that~\cite{we2018seim, we2018adbis, thepaper, we2018beyondmr, trofimov2018consistency, webirte, we2019debs}. Chapter~\ref{thesis-chapter-substreams-consistency} and Chapter~\ref{thesis-chapter-tracker} are based on papers on \tracker\ framework and substream management~\cite{10.1145/3524860.3539809, trofimov2023bounding}.

The complete list of published material along with assigned contributions is the following:

\begin{enumerate}
    \item Artem Trofimov, Nikita Sokolov, Nikita Marshalkin, Igor Kuralenok, and Boris Novikov. \textbf{Bounding substreams in distributed stream processing}. Information Systems, Volume 117, 2023. \newline
    
    \textit{Contribution:} implementation of the main parts of the substream management framework and the formal model for the substream management domain, leading the writing of the paper.

    \item Artem Trofimov, Nikita Sokolov, Nikita Marshalkin, Igor Kuralenok, and Boris Novikov. \textbf{Substream management in distributed streaming dataflows}. In Proceedings of the 16th ACM International Conference on Distributed and Event-Based Systems (DEBS '22), pp. 55-66. ACM, 2022. \newline
    
    \textit{Contribution:} implementation of the main parts of the substream management framework and the formal model for the substream management domain, leading the writing of the paper.

    \item Igor Kuralenok, Artem Trofimov, Nikita Marshalkin, and Boris Novikov. \textbf{Deterministic Model for Distributed Speculative Stream Processing}. In European Conference on Advances in Databases and Information Systems (pp. 233-246). Cham: Springer International Publishing, 2018. \newline
    
    \textit{Contribution:} refinement of the mechanism for ensuring determinism, conducting the experiments, co-authoring the paper.

    \item Igor Kuralenok, Nikita Marshalkin, Artem Trofimov, and Boris Novikov. \textbf{An optimistic approach to handle out-of-order events within analytical stream processing}. In Third Conference on Software Engineering and Information Management (SEIM-2018) (pp. 22-29), 2018. \newline
    
    \textit{Contribution:} participation in implementation of the mechanism for handling out-of-order elements, conducting the experiments, co-authoring the paper.

    \item Artem Trofimov, Igor Kuralenok, Nikita Marshalkin, and Boris Novikov. \textbf{Delivery, consistency, and determinism: rethinking guarantees in distributed stream processing}. arXiv preprint arXiv:1907.06250, 2019. \newline
    
    \textit{Contribution:} implementation of the main parts of the formal model for delivery guarantees and the mechanism for achieving exactly-once, leading the writing of the paper.

    \item Artem Trofimov, Nikita Sokolov, Mikhail Shavkunov, Igor Kuralenok, and Boris Novikov. \textbf{Distributed Classification of Text Streams: Limitations, Challenges, and Solutions}. In Proceedings of Real-Time Business Intelligence and Analytics (pp. 1-6), 2019. \newline
    
    \textit{Contribution:} problem setup, conducting the experiments, leading the writing of the paper.

    \item Artem Trofimov. \textbf{Consistency Maintenance in Distributed Analytical Stream Processing}. In New Trends in Databases and Information Systems: ADBIS 2018 Short Papers and Workshops, AI* QA, BIGPMED, CSACDB, M2U, BigDataMAPS, ISTREND, DC, Budapest, Hungary, September, 2-5, 2018, Proceedings 22. Springer International Publishing, 2018. \newline
    
    \textit{Contribution:} leading the writing of the paper.

    \item Igor Kuralenok, Artem Trofimov, Nikita Marshalkin, and Boris Novikov. \textbf{Flamestream: model and runtime for distributed stream processing}. In Proceedings of the 5th ACM SIGMOD Workshop on Algorithms and Systems for MapReduce and Beyond (pp. 1-2), 2018 \newline
    
    \textit{Contribution:} conducting the experiments, co-authoring the paper.

    \item Artem Trofimov, Mikhail Shavkunov, Sergey Reznick, Nikita Sokolov, Mikhail Yutman, Igor Kuralenok, and Boris Novikov. \textbf{Reproducible and reliable distributed classification of text streams}. In Proceedings of the 13th ACM International Conference on Distributed and Event-based Systems (pp. 264-265), 2019. \newline
    
    \textit{Contribution:} problem setup, leading the writing of the paper.
\end{enumerate}