\documentclass{vldb}

\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{hyperref}
\usepackage{algorithm} % for algorithms
% \usepackage{algorithmicx}
% \usepackage{algorithm2e} % for algorithms
\usepackage{algpseudocode}
% \usepackage{booktabs} % For formal tables
\algdef{SE}[SUBALG]{Indent}{EndIndent}{}{\algorithmicend\ }%
\algtext*{Indent}
\algtext*{EndIndent}

\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[flushleft]{threeparttable}
\usetikzlibrary{backgrounds, calc, positioning, fit, decorations.pathreplacing}
% \theoremstyle{remark}

\pagestyle{empty} % removes running headers

\newcommand{\PicScale}{0.5}
\newcommand {\FlameStream} {FlameStream}
\newcommand {\tracker} {Tracker}
\newcommand {\acker} {Acker}

\newtheorem{lemma}{Lemma}

% Include information below and uncomment for camera ready
\vldbTitle{Substreams Management in Distributed Streaming Dataflows}
\vldbAuthors{Artem Trofimov, Nikita Sokolov, Igor Kuralenok, Nikita Marshalkin, and Boris Novikov}
\vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
\vldbVolume{12}
\vldbNumber{xxx}
\vldbYear{2019}

\begin{document}

\title {Substream Management in Distributed Streaming Dataflows}

\numberofauthors{5}

\author{
\alignauthor
Artem Trofimov\\
    \affaddr{Yandex}\\
    \affaddr{Saint Petersburg, Russia}\\
    \email{tomato@yandex-team.ru}
\alignauthor
Nikita Sokolov\\
    \affaddr{Yandex}\\
    \affaddr{Saint Petersburg, Russia}\\
    \email{faucct@yandex-team.ru}
\alignauthor
Igor Kuralenok\\
    \affaddr{Yandex}\\
    \affaddr{Saint Petersburg, Russia}\\
    \email{solar@yandex-team.ru}
\and 
\alignauthor
Nikita Marshalkin\\
    \affaddr{VK}\\
    \affaddr{Saint Petersburg, Russia}\\
    \email{n.marshalkin@corp.vk.com}
\alignauthor
Boris Novikov\\
    \affaddr{HSE University}\\
    \affaddr{Saint Petersburg, Russia}\\
    \email{borisnov@acm.org}
}

\maketitle

\begin{abstract}
Stream Processing Engines (SPEs) handle a potentially infinite sequence of data elements. This sequence often can be viewed as a mixture of finite substreams with distinct properties: event times ranges, payload values, etc. Some operators ought to know since then there will be no longer elements with specified properties, i.e., when the substream ends. For instance, stateful streaming operators should clear outdated state to avoid running out of memory. Time window operators should release output after all elements within the specified time range arrived. Most state-of-the-art SPEs use {\em punctuations} to manage substreams. The punctuations approach is powerful but has limitations: it does not support cyclic dataflows, poorly scalable due to frequent broadcasts, and is inefficient for small substreams because of high network traffic overhead. In this work, we present a new substream management framework called \tracker\ that supports cyclic dataflows and tiny substreams (even substreams containing a single element). This method bases on the idea that a special agent within an SPE monitors for substreams and notifies SPE nodes about substreams lifespan events. We demonstrate that our implementation can provide the same semantics of notifications as punctuations but implies lower network traffic overhead. Experiments show that our technique is scalable, efficiently handles real-world cyclic dataflows, and outperforms punctuations for small substreams.
\end{abstract}

% \keywords{Data streams, exactly-once, drifting state, optimistic OOP}

\thispagestyle{empty}

\section {Introduction}
\input {fs-acker-intro}

\section{Substreams management}
\input {fs-acker-preliminaries}

\section{Tracker framework}
\input {fs-acker-overview}

\section{Tracker Implementation}
\input {fs-acker-impl}

\section {Experiments}
\input {fs-acker-experiments.tex}

\section{Related Work}
\input {fs-acker-related.tex}

\section {Conclusion}
\input {fs-acker-conclusion.tex}

\appendix 
\section {Punctuations} \label{appendix:punctuations-proof}
\section {\tracker} \label{appendix:tracker-proof}

\bibliographystyle{abbrvurl}
% \bibliography{bibliography/flame-stream}
\bibliography{../../bibliography/flame-stream}

\end {document}

\endinput
